---
title: "Project Code"
author: "Diego Brule, Nicolas Tamara"
date: "27/11/2024"
output: github_document
---

```{r}

library(dplyr)
library(ggplot2)
library(ggrepel)
library(readxl)
library(DIMORA)
library(lmtest)

```


```{r}

#All_data <- read_xlsx('Product Data finale.xlsx')
All_data <- read.csv('Product_Data Q3_2022 to Q4_2023.csv')

```

```{r}

Date_Converter <- function(month, year){
  Month_name <- c('January', 'February', 'March', 'April', 'May', 'June', 'July', 'August', 'September', 'October', 'November', 'December')
  paste(month, year)
}

Date_Converter(3, 2023)

```

#Glimpse at the data

First, let's look at the different time series for the different medicine brands

```{r}

#Creating a column 'mes' to represent total months. i.e. January 2023 is 13 instead of 1. This is to facilitate graphing the time series.

#Creating a column 'Month' to have a more interpretable x axis. Can't figure out how to work this yet.

All_data <- All_data %>% mutate(Mes = ifelse(AnnioCorte == 2022, MesFactura, MesFactura + 12 ))

#All_data <- All_data %>% mutate(Month = Date_Converter(MesFactura, AnnioCorte))

#ggplot(All_data, aes(x=Month, y=TotalUnidades, group = DescripcionComercial, colour = DescripcionComercial)) + geom_line() + scale_x_discrete(guide=guide_axis(check.overlap = TRUE))

ggplot(All_data, aes(x=Mes, y=TotalUnidades, group = DescripcionComercial, colour = DescripcionComercial)) + geom_line()

```

We see that there are three brands: Iclomod, Lebrina, Gilenya That dominate the market in terms of units sold. Let's zoom in on those, and the other lower-selling brands separately.

```{r}

#Separate Data by different drug brands to ease analysis, select only ts variables

Aximod <- All_data %>% filter(DescripcionComercial == 'AXIMOD')
Fincler <- All_data %>% filter(DescripcionComercial == 'FINCLER')
Gilenya <- All_data %>% filter(DescripcionComercial == 'GILENYA')
Iclomod <- All_data %>% filter(DescripcionComercial == 'ICLOMOD')
Kesimpta <- All_data %>% filter(DescripcionComercial == 'KESIMPTA')
Lebrina <- All_data %>% filter(DescripcionComercial == 'LEBRINA')
Lemtrada <- All_data %>% filter(DescripcionComercial == 'LEMTRADA')
Limostad <- All_data %>% filter(DescripcionComercial == 'LIMOSTAD')
Mavenclad <- All_data %>% filter(DescripcionComercial == 'MAVENCLAD')
Ocrevus <- All_data %>% filter(DescripcionComercial == 'OCREVUS')
Singomod <- All_data %>% filter(DescripcionComercial == 'SINGOMOD')
Tysabri <- All_data %>% filter(DescripcionComercial == 'TYSABRI')


```

```{r}

Trend_Data <- All_data %>% filter((DescripcionComercial == 'GILENYA') | (DescripcionComercial == 'LEBRINA') | (DescripcionComercial == 'ICLOMOD'))

ggplot(Trend_Data, aes(x=Mes, y=TotalUnidades, group = DescripcionComercial, colour = DescripcionComercial)) + geom_line()

```

```{r}

Flat_Data <- All_data %>% filter((DescripcionComercial != 'GILENYA') & (DescripcionComercial != 'LEBRINA') & (DescripcionComercial != 'ICLOMOD'))

ggplot(Flat_Data, aes(x=Mes, y=TotalUnidades, group = DescripcionComercial, colour = DescripcionComercial)) + geom_line()

```

#First round of model fitting

Since the top three brands have much higer volumes of units sold, the movements and fluctuations in sales are also bigger, so we will start by fitting a simple bass model, while we will start with a simple linear model for the other brands.


```{r}
BM_Iclomod <- BM(Iclomod$TotalUnidades, display = T)
summary(BM_Iclomod)

ggplot(Iclomod, aes(x=Mes, y=TotalUnidades)) + geom_point() + geom_line(linetype='dashed') + geom_line(y=make.instantaneous(BM_Iclomod$fitted), color='red') + ggtitle('Iclomod Units sold and BM fit')

acf(BM_Iclomod$residuals)

ggplot(Iclomod, aes(x=Mes, y=BM_Iclomod$residuals)) + geom_point() + geom_hline(yintercept = 0, linetype = 'dashed') + ggtitle('Iclomod residuals of simple BM') + labs(x='Month', y='Residuals')

```

At a shallow level, the BM fit for Iclomod data is good, it has high $R^2$ and a relevant p-value for its variables. However, analysing the residuals graph we can see that while it contains a more or less constant variance, which is a good sign, it seems to have an underlying pattern, which means there is a pattern in the data that was not captured in the model. (Insert next model idea)

```{r}

BM_Gilenya <- BM(Gilenya$TotalUnidades)

summary(BM_Gilenya)

ggplot(Gilenya, aes(x=Mes, y=TotalUnidades)) + geom_point() + geom_line(linetype='dashed') + geom_line(y=make.instantaneous(BM_Gilenya$fitted), color='red') + ggtitle('Gilenya units sold and simple BM fit')

acf(BM_Gilenya$residuals)

ggplot(Gilenya, aes(x=Mes, y=BM_Gilenya$residuals)) + geom_point() + geom_hline(yintercept = 0, linetype = 'dashed') + ggtitle('Gilenya residuals for simple BM fit')

```

Similarly to the BM fit for Iclomod, we have a high $R^2$ and relevant p-values, and constant variance in the residuals, but there seems to be an underlying pattern in the residuals. 


```{r}

BM_Lebrina <- BM(Lebrina$TotalUnidades, display = T)

summary(BM_Lebrina)

ggplot(Lebrina, aes(x=Mes, y=TotalUnidades)) + geom_point() + geom_line(linetype='dashed') + geom_line(y=make.instantaneous(BM_Lebrina$fitted), color='red') + ggtitle('Lebrina units sold and simple BM fit')

acf(BM_Lebrina$residuals)

ggplot(Lebrina, aes(x=Mes, y=BM_Lebrina$residuals)) + geom_point() + geom_hline(yintercept = 0, linetype = 'dashed') + ggtitle('Lebrina residuals for simple BM fit')

```

And again, we see a high $R^2$, relevant p-values but less constant variance and seemingly an underlying residuals pattern.

```{r}

# A simple linear model for the smaller selling medications

lm_Kesimpta <- lm(Kesimpta$TotalUnidades ~ Kesimpta$Mes)
acf(Kesimpta$TotalUnidades)

summary(lm_Kesimpta)

ggplot(Kesimpta, aes(x=Mes, y=TotalUnidades)) + geom_line(linetype='dashed') + geom_point() +geom_abline(intercept = lm_Kesimpta$coefficients[1], slope = lm_Kesimpta$coefficients[2], color='red')

dwtest(lm_Kesimpta)

ggplot(Kesimpta, aes(x=Mes, y=lm_Kesimpta$residuals)) + geom_point() + geom_hline(yintercept = 0, linetype='dashed')

acf(lm_Kesimpta$residuals)

```

Kesimpta is one of the lower-selling brands, so we first fit a linear model to it. The model is good, it has high $R^2$ and relevant estimators according to their p-values, but as time goes on the variance grows. This is a sign of something missing in the model

```{r}

#Simple Linear model for Axmimod

lm_Aximod <- lm(Aximod$TotalUnidades ~ Aximod$Mes)
acf(Aximod$TotalUnidades)

summary(lm_Aximod)

ggplot(Aximod, aes(x=Mes, y=TotalUnidades)) + geom_line(linetype='dashed') + geom_point() +geom_abline(intercept = lm_Aximod$coefficients[1], slope = lm_Aximod$coefficients[2], color='red')

dwtest(lm_Aximod)

ggplot(Aximod, aes(x=Mes, y=lm_Aximod$residuals)) + geom_point() + geom_hline(yintercept = 0, linetype='dashed')

acf(lm_Aximod$residuals)

```

```{r}

#Simple linear model for Fincler

lm_Fincler <- lm(Fincler$TotalUnidades ~ Fincler$Mes)
acf(Fincler$TotalUnidades)

summary(lm_Fincler)

ggplot(Fincler, aes(x=Mes, y=TotalUnidades)) + geom_line(linetype='dashed') + geom_point() +geom_abline(intercept = lm_Fincler$coefficients[1], slope = lm_Fincler$coefficients[2], color='red')

dwtest(lm_Fincler)

ggplot(Fincler, aes(x=Mes, y=lm_Fincler$residuals)) + geom_point() + geom_hline(yintercept = 0, linetype='dashed')

acf(lm_Fincler$residuals)

```

```{r}

lm_Lemtrada <- lm(Lemtrada$TotalUnidades ~ Lemtrada$Mes)
acf(Lemtrada$TotalUnidades)

summary(lm_Lemtrada)

ggplot(Lemtrada, aes(x=Mes, y=TotalUnidades)) + geom_line(linetype='dashed') + geom_point() +geom_abline(intercept = lm_Lemtrada$coefficients[1], slope = lm_Lemtrada$coefficients[2], color='red')

dwtest(lm_Lemtrada)

ggplot(Lemtrada, aes(x=Mes, y=lm_Lemtrada$residuals)) + geom_point() + geom_hline(yintercept = 0, linetype='dashed')

acf(lm_Lemtrada$residuals)

```

```{r}

lm_Limostad <- lm(Limostad$TotalUnidades ~ Limostad$Mes)
acf(Limostad$TotalUnidades)

summary(lm_Limostad)

ggplot(Limostad, aes(x=Mes, y=TotalUnidades)) + geom_line(linetype='dashed') + geom_point() +geom_abline(intercept = lm_Limostad$coefficients[1], slope = lm_Limostad$coefficients[2], color='red')

dwtest(lm_Limostad)

ggplot(Limostad, aes(x=Mes, y=lm_Limostad$residuals)) + geom_point() + geom_hline(yintercept = 0, linetype='dashed')

acf(lm_Limostad$residuals)

```
```{r}

lm_Mavenclad <- lm(Mavenclad$TotalUnidades ~ Mavenclad$Mes)

acf(Mavenclad$TotalUnidades)

summary(lm_Lemtrada)

ggplot(Mavenclad, aes(x=Mes, y=TotalUnidades)) + geom_line(linetype='dashed') + geom_point() +geom_abline(intercept = lm_Mavenclad$coefficients[1], slope = lm_Mavenclad$coefficients[2], color='red')

dwtest(lm_Mavenclad)

ggplot(Mavenclad, aes(x=Mes, y=lm_Mavenclad$residuals)) + geom_point() + geom_hline(yintercept = 0, linetype='dashed')

acf(lm_Mavenclad$residuals)

```
```{r}

lm_Ocrevus <- lm(Ocrevus$TotalUnidades ~ Ocrevus$Mes)
acf(Ocrevus$TotalUnidades)

summary(lm_Ocrevus)

ggplot(Ocrevus, aes(x=Mes, y=TotalUnidades)) + geom_line(linetype='dashed') + geom_point() +geom_abline(intercept = lm_Ocrevus$coefficients[1], slope = lm_Ocrevus$coefficients[2], color='red')

dwtest(lm_Ocrevus)

ggplot(Ocrevus, aes(x=Mes, y=lm_Ocrevus$residuals)) + geom_point() + geom_hline(yintercept = 0, linetype='dashed')

acf(lm_Ocrevus$residuals)

```
```{r}

lm_Singomod <- lm(Singomod$TotalUnidades ~ Singomod$Mes)
acf(Singomod$TotalUnidades)

summary(lm_Singomod)

ggplot(Singomod, aes(x=Mes, y=TotalUnidades)) + geom_line(linetype='dashed') + geom_point() +geom_abline(intercept = lm_Singomod$coefficients[1], slope = lm_Singomod$coefficients[2], color='red')

dwtest(lm_Singomod)

ggplot(Singomod, aes(x=Mes, y=lm_Singomod$residuals)) + geom_point() + geom_hline(yintercept = 0, linetype='dashed')

acf(lm_Singomod$residuals)

```
```{r}

lm_Tysabri <- lm(Tysabri$TotalUnidades ~ Tysabri$Mes)
acf(Tysabri$TotalUnidades)

summary(lm_Tysabri)

ggplot(Tysabri, aes(x=Mes, y=TotalUnidades)) + geom_line(linetype='dashed') + geom_point() +geom_abline(intercept = lm_Tysabri$coefficients[1], slope = lm_Tysabri$coefficients[2], color='red')

dwtest(lm_Tysabri)

ggplot(Tysabri, aes(x=Mes, y=lm_Tysabri$residuals)) + geom_point() + geom_hline(yintercept = 0, linetype='dashed')

acf(lm_Tysabri$residuals)

```
